Rhiredis
========

Rhiredis is a wrapper of hiredis, which is a minimalistic C client library for the Redis database.

Rhiredis is much faster than rredis on CRAN. However, it internally converts every objects to `character` and send them to redis server. The user could still save R objects into redis server with `serialize` and `base64enc::base64encode`. Please see the benchmark for example.

## Dependencies

- boost shared_ptr
- hiredis
- R packages
    - Rcpp
    - bit64

## Install

```sh
git clone https://github.com/wush978/Rhiredis.git
cd Rhiredis
git submodule init 
git submodule update --recursive
R CMD INSTALL .
```

## Getting Start

```r
library(Rhiredis)

con <- redisConnect()
redisSET("a", "b") # "OK"
redisGET("a")
```

The name pattern of functions is `redis<Command name>`. Please read http://redis.io/commands for more details about commands of redis.

## Benchmark

### Character

The following example compares the time of inserting strings to redis server between `Rhiredis` and `rredis`.

```{r benchmark-character}
suppressPackageStartupMessages(library(Rhiredis))
suppressPackageStartupMessages(library(base64enc))
suppressPackageStartupMessages(library(rredis))
suppressPackageStartupMessages(library(rbenchmark))

con <- Rhiredis::redisConnect()
rredis::redisConnect()

benchmark(
	Rhiredis::redisSET("a1", "b1"),
	rredis::redisSet("a2", "b2")
)

Rhiredis::redisGET("a1")
rredis::redisGet("a2")
```

### Object

The following example compares the time of inserting R object into redis server between `Rhiredis` and `rredis`.

```{r benchmark-object}
suppressPackageStartupMessages(library(Rhiredis))
suppressPackageStartupMessages(library(base64enc))
suppressPackageStartupMessages(library(rredis))
suppressPackageStartupMessages(library(rbenchmark))

con <- Rhiredis::redisConnect()
rredis::redisConnect()

benchmark(
	Rhiredis::redisSET("a1", base64encode(serialize(iris, NULL))),
	rredis::redisSet("a2", iris)
)

all.equal(unserialize(base64decode(Rhiredis::redisGET("a1"))), iris)
all.equal(rredis::redisGet("a2"), iris)
```

## License

GNU GENERAL PUBLIC LICENSE Version 3, 29 June 2007
