---
title: A Brief Introduction to Redis

# Use letters for affiliations
author:
  - name: Dirk Eddelbuettel
    affiliation: 1
address:
  - code: 1
    address: Department of Statistics, University of Illinois, Urbana-Champaign, IL, USA

# For footer text  TODO(fold into template, allow free form two-authors)
lead_author_surname: Eddelbuettel

# Place DOI URL or CRAN Package URL here
doi: "https://cran.r-project.org/package=RcppRedis"

# Abstract
abstract: |
  This note provides a brief introduction to Redis highlighting its
  usefulness in multi-lingual statistical computing.

# Font size of the document, values of 9pt (default), 10pt, 11pt and 12pt
fontsize: 9pt

# Optional: Force one-column layout, default is two-column
two_column: true

# Optional: Enables lineno mode, but only if one_column mode is also true
#lineno: true

# Optional: Enable one-sided layout, default is two-sided
#one_sided: true

# Optional: Enable section numbering, default is unnumbered
#numbersections: true

# Optional: Specify the depth of section number, default is 5
#secnumdepth: 5

# Optional: Skip inserting final break between acknowledgements, default is false
skip_final_break: true

# Optional: Bibliography
bibliography: redis

# Optional: Enable a 'Draft' watermark on the document
watermark: false

# Customize footer, eg by referencing the vignette
footer_contents: "Redis Intro"

# Produce a pinp document
output:
  pinp::pinp:
    collapse: true
    keep_tex: false

header-includes: >
  \newcommand{\proglang}[1]{\textsf{#1}}
  \newcommand{\pkg}[1]{\textbf{#1}}
  \newcommand{\R}{\proglang{R}}
  \newcommand{\Py}{\proglang{Python}}
  \newcommand{\Redis}{\proglang{Redis}}

# Required: Vignette metadata for inclusion in a package.
vignette: >
  %\VignetteIndexEntry{Brief Redis Introduction}
  %\VignettePackage{RcppRedis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\usepackage[utf8]{inputenc}
---

```{r initialsetup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE)
```


# Overview and Introduction

[Redis][redis] \citep{Redis} is a very popular, very powerful, and very widely-used 'in-memory database-structure
store'. It runs as a background process (a "daemon" in Unix lingo) and can be accessed locally or
across a network making it very popular choice for 'data caches'. There is more to say about
[Redis][redis] than we possibly could in a _short_ vignette introducing it, and other places already
do so. The [Little Redis Book](https://www.openmymind.net/2012/1/23/The-Little-Redis-Book/) \citep{Seguin:2021:Redis}, while a
decade old (!!) is a fabulous _short_ start.  The [official site][redis] is very good as well (but
by now a little overwhelming as so many features have been added).

This vignette aims highlight two aspects: how easy it is to use [Redis][redis] on simple examples,
and also to stress how [Redis][redis] enables easy _multi-lingual_ computing as it can act as a
'middle-man' between any set of languages capable of speaking the [Redis][redis] protocol -- which
may cover most if not all common languages one may want to use!

# Data Structure Example One: Key-Value Setter and Getter

We will show several simple examples for the

- `redis-cli` command used directly or via shell scripts
- \Py via the standard \Py package for [Redis][redis]
- \proglang{C} / \proglang{C++} via the [hiredis][hiredis] library
- \R via \pkg{RcppRedis} \citep{CRAN:RcppRedis} utilising the [hiredis][hiredis] library

to demonstrate how different languages all can write to and read from [Redis][redis].  Our first example will
use the simplest possibly data structure, a simple `SET` and `GET` of a key-value pair.

## Command-Line

`redis-cli` is command-line client.  Use is straightforward as shown an simply key-value getter and
setter. We show use in 'shell script' form here, but the same commands also work interactively.

```{bash cliGetSet}
## note that document processing will show all
## three results at once as opposed to one at time
redis-cli SET ice-cream chocolate
redis-cli GET ice-cream
redis-cli GET ice-cream
```

Here, as in general, we will omit hostname and authentication arguments: on the same machine,
`redis-cli` and the background `redis` process should work 'as is'. For access across a (local or
remote) network, the configuration will have to be altered to permit access at given network
interfaces and IP address ranges.

We show the [redis][redis] commands used in uppercase notation, this is in line with the
documentation. Note, however, that [Redis][redis] itself is case-insensitive here so `set` is equivalent to
`SET`.

## Python

[Redis][redis] does have bindings for most, if not all, languages to
computing with data. Here is a simple \proglang{Python} example.

```{python pyGetSet}
import redis

redishost = "localhost"
redisserver = redis.StrictRedis(redishost)

key = "ice-cream"
val = "strawberry"
res = redisserver.set(key, val)
print("Set", val, "under", key, "with result", res)

key = "ice-cream"
val = redisserver.get(key)
print("Got", val, "from", key)
```

For \Py, the redis commands are generally mapped to (lower-case named) member functions of the
instantiated redis connection object, here `redisserver`.

## C / C++

Compiled languages work similarly. For \proglang{C} and \proglang{C++}, the [hiredis][hiredis] 'minimalistic' library from
the [Redis][redis] project can be used---as it is by \pkg{RcppRedis}. Here we only show the code without executing
it. This example is included in the package are as the preceding ones. \proglang{C} and \proglang{C++} work similarly to
the interactive or \Py commands. A simplified (and incomplete, see the `examples/`
directory of the package for more) example of writing to [Redis][redis] would be

```c++
redisContext *prc;    // pointer to redis context

std::string host = "127.0.0.1";
int port = 6379;

prc = redisConnect(host.c_str(), port);
// should check error here
redisReply *reply = (redisReply*)
    redisCommand(prc, "SET ice-cream %s", value);
// should check reply here
```

Reading is done by submitting for example a `GET` command for a key after which the `redisReply`
contains the reply string.

## R

The \pkg{RcppRedis} packages uses \pkg{Rcpp} Modules along with \pkg{Rcpp}
\citep{CRAN:Rcpp,TAS:Rcpp} to connect the [hiredis][hiredis] library to \R.  A simple \R example
follows.

```{r RexSetGet}
library(RcppRedis)
redis <- new(Redis, "localhost")
redis$set("ice-cream", "hazelnut")
redis$get("ice-cream")
```

# Data Structure Example Two: Hashes

[Redis][redis] has support for a number of standard data structures including hashes.
The official documentation list [sixteen commands](https://redis.io/commands#hash) in the
corresponding group covering writing (`hset`), reading (`hget`), checking for key (`hexists`),
deleting a key (`hdel`) and more.

```{bash cliHash}
redis-cli HSET myhash abc 42
redis-cli HSET myhash def "some text"
```

We can illustrate reading and checking from Python:

```{python pyHash}
print(redisserver.hget("myhash", "abc"))
print(redisserver.hget("myhash", "def"))
print(redisserver.hexists("myhash", "xyz"))
```

For historical reasons, \pkg{RcppRedis} converts to/from \proglang{R} internal serialization on
hash operations so it cannot _directly_ interoperate with these examples as they not deploy
\R-internal representation. The package has however a 'catch-all' command `exec` (which
excutes a given \Redis command string) which can be used here:

```{r rHash}
redis$exec("HGET myhash abc")
redis$exec("HGET myhash def")
redis$exec("HEXISTS myhash xyz")
```

# Data Structure Example Three: Sets

Sets are another basic data structure that is well-understood. In sets, elements can be added (once),
removed (if present), and sets can be joined, intersected and differenced.

```{bash cliSet}
redis-cli SADD myset puppy
redis-cli SADD myset kitten
redis-cli SADD otherset birdie
redis-cli SADD otherset kitten
redis-cli SINTER myset otherset
```

We can do the same in \Py. Here we showi only the final intersect command---the set-addition
commands are equivalent across implementations as are most other [Redis] command.

```{python, pySetPre, include=FALSE}
import redis
redisserver = redis.StrictRedis("localhost")
```
```{python, pySet}
print(redisserver.sinter("myset", "otherset"))
```

And similarly in \R where `exec` returns a list:

```{r rSet}
redis$exec("SINTER myset otherset")
```

# Data Structure Example Four: Lists

So far we have looked at setters and getters for values, hashes, and sets. All of these
covered only one value per key. But \Redis has support for many more types.

```{bash cliList}
redis-cli LPUSH mylist chocolate
redis-cli LPUSH mylist strawberry vanilla
redis-cli LLEN mylist
```

We can access the list in \Py. Here we show access by index. Note that the index is zero-based, so
'one' accesses the middle element in a list of length three.


```{python, pyList}
print(redisserver.lindex("mylist", 1))
```

In \R, using the 'list range' command for elements 0 to 1:

```{r rList}
redis$exec("LRANGE mylist 0 1")
```

The \pkg{RcppRedis} list commands (under the 'standard' names) work on serialized R objects so we
once again utilize the `exec` command to execute this using the 'standard' name. As access to
unserialized data is useful, the package also two alternates for numeric and string return data:

```{r rList2}
redis$listRangeAsStrings("mylist", 0, 1)
```

# Data Structure Example Five: Sorted Sets

[Redis][redis] offers another data structure that can be of interest to us for use in time series.
Recall how packages \pkg{zoo} \citep{CRAN:zoo} and \pkg{xts} \citep{CRAN:xts} are, essentially,
indexed containers around (numeric) matrices with a sorting index. This is commonly the `Date` type
in \R for daily data, or a `POSIXct` datimetime type for intra-daily data at approximately a
microsecond resolution. One can then index by day or datetime, subset, merge, ...  We can store such
data in \Redis using sorted sets using the index as the first column.  A quick \R example
illustrates.

```{r sortedSet1}
m1 <- matrix(c(100, 1, 2, 3), 1)
redis$zadd("myz", m1) # add m1 indexed at 100
m2 <- matrix(c(105, 2, 2, 4), 1)
redis$zadd("myz", m2) # add m1 indexed at 105
```

In this first example we insert two matrices (with three values each) index at 100 and 105,
respectively, to the sorted set under key `myz`. We will then ask for a range of data over the range
from 90 to 120 which will include both sets of observations.

```{r sortedSet2}
res <- redis$zrangebyscore("myz", 90, 120)
res
```


# Communication Example: Publish/Subscribe

We have seen above that writen a _value_ to a particular _key_ into a list, set, or sorted set is
straightforward.  So is publishing into a _channel_. \Redis keeps track of the current subscribers
to a channel and dispatches the published content.

Subscribers can join, and leave, anytime. Data is accessible via the publish/subscribe (or
"pub/sub") mechanism while being subscribe. There is no mechanism _within pub/sub_ to obtain 'older'
values, or to re-request values. Such services can however be provided by \Redis is its capacity of
a data store.

As subscription is typically blocking, we cannnot show a simple example in the vignette. But an
illustration (without running code) follows.

```r
ch1 <- function(x) { cat("Got", x, "in ch1\n") }
redis$subscribe("ch1")
```

Here we declare a callback function which by our convention uses the same name as the channel. So in
the next when the subscription is activated, the callback function is registered with the current
\Redis object. Once another process or entity publishes to the channel, the callback function will
be invoked along with the value published on the channel.

# Application Example: Hash R Objects

The ability to serialize R object makes it particularly to store R objects directly.  This can be
useful for data sets, and well as generated data

```{r treesEx}
fit <- lm(Volume ~ . - 1, data=trees)
redis$hset("myhash", "data", trees)
redis$hset("myhash", "fit", fit)
fit2 <- redis$hget("myhash", "fit")
all.equal(fit, fit2)
```
The retrieved model fit is equal to the one we stored in [Redis]. We can also re-fit on the
retrieved data and obtain the same coefficient.  (The fit object stores information about the data
set which differs here for technical reason internal to \R; the values are the same.)

```{r treesEx2}
data2 <- redis$hget("myhash", "data")
fit2 <- redis$hget("myhash", "fit")
fit3 <- lm(Volume ~ . - 1, data=data2)
all.equal(coef(fit2), coef(fit3))
```

# Summary

This vignettet introduces the [Redis][redis] data structure engine, and demonstrates how reading
and writing different data types from different programming languages including \R, \Py and shell
is concise and effective.  A final example of storing an \R dataset and model fit further
illustrates the versatility of \Redis.

<!--# Cleanup-->
```{bash cleanup, include=FALSE}
redis-cli del ice-cream
redis-cli hdel myhash abc
redis-cli hdel myhash def
redis-cli del mylist
redis-cli del myset
redis-cli del otherset
redis-cli del myz
```

[redis]: https://redis.io
[hiredis]: https://github.com/redis/hiredis
